Feasibility of Harvesting Power To Run A Domestic Water Meter Using Streaming Cell Technology

### 2014-10-14 paper.tex
Kyle's proofread

    Figure~\ref{fig:headloss} shows the pressure head loss curve from a water meter typically installed at New Zealand homes (Kent PSMT 25mm).~\cite{WatercareNewZealand2014}
    Using this curve we calculate power dissipation in a water meter during a washing machine cycle, shower, and toilet flush; presented as Fig.~\ref{fig:powerDissipated_meter}.
    The total energy dissipated within the meter for each events is:
    \begin{itemize}
    \item \SI{547}{\joule} per load of washing,
    \item \SI{222}{\joule} per shower, and
    \item \SI{24.3}{\joule} per flush of the toilet.
    \end{itemize}

@@

    \section{Estimation of Energy Requirements}
    \label{sect:powerRequirements}
    In \cite{Jones2011} we measure the energy efficiency of various low-power microcontrollers.
    Here we take measured data from that paper to estimate a smart meter's energy requirements.
    Measurements of an Atmel ATtiny 25V have been used as it offered good performance over a wide range of processor functions.
    A crude estimation of the processor's event loop is as follows:
    \begin{itemize}
    \item Sleep for \SI{1}{\second} (\SI{97.4}{\nano\joule})
    \item Execute 1 000 instructions (\SI{1.14}{\micro\joule})
    \item Take 2 ADC measurements (\SI{2.56}{\nano\joule})
    \item Write 2 bytes to non-volatile memory (\SI{79.0}{\micro\joule})
    \end{itemize}
    Every six hours the device will transmit metered data by doing the following actions:
    \begin{itemize}
    \item Execute 1 000 000 instructions (\SI{1.14}{\milli\joule})
    \item Transmit 100 bytes of data over ZigBee radio (\SI{12.3}{\milli\joule})
    \item Write 10 bytes to non-volatile memory (\SI{394}{\micro\joule})
    \end{itemize}

@@

    To increase the efficiency of the channel the internal height needs to be reduced by a factor of \SI{52 600}, from \SI{71}{\micro\metre} to \SI{1.35}{\nano\metre} (the midpoint between \SI{0.6}{\nano\metre} and \SI{2.1}{\nano\metre}).
    Doing so decreases the normalised output power of the streaming cell to \SI{2.89}{\pico\watt\per\metre}.
    Therefore the width of a channel required to generate the target \SI{5.6}{\milli\watt} is \SI{1.94}{\giga\metre}.

@@

    \section{Conclusion}
    \label{sect:conclusion}
    Approximately $100\times$ the required power to operate an automatic meter reader and transmitter is already dissipated inside today's mechanical water meters.
    The streaming cells we constructed produced conversion efficiencies in the order of \SI{1}{\micro\percent}.
    Scaling of these streaming cells to create a viable harvester would result in a \SI{22}{\cubic\metre} volume of glass with a mass of \SI{54}{\mega\gram}.
    This is far from being a practical solution to the battery operated devices currently in use.

>>>>

    Figure~\ref{fig:headloss} shows the pressure head loss curve from a water meter typically installed at New Zealand homes (Kent PSMT \SI{25}{\milli\meter}).~\cite{WatercareNewZealand2014}
    Using this curve we calculate power dissipation in a water meter during a washing machine cycle, shower, and toilet flush; presented as Fig.~\ref{fig:powerDissipated_meter}.
    The total energy dissipated within the meter for each events is:
    \begin{itemize}
    \item \SI{547}{\joule} per load of washing,
    \item \SI{222}{\joule} per shower, and
    \item \SI{24.3}{\joule} per flush of the toilet.
    \end{itemize}

@@

    \section{Estimation of Energy Requirements}
    \label{sect:powerRequirements}
    In \cite{Jones2011} we measure the energy efficiency of various low-power microcontrollers.
    Here we take measured data from that paper to estimate a smart meter's energy requirements.
    Measurements of an Atmel ATtiny 25V have been used as it offered good performance over a wide range of processor functions.
    A crude estimation of the processor's event loop is as follows:
    \begin{itemize}
    \item Sleep for \SI{1}{\second} (\SI{97.4}{\nano\joule})
    \item Execute \SI{1000} instructions (\SI{1.14}{\micro\joule})
    \item Take 2 ADC measurements (\SI{2.56}{\nano\joule})
    \item Write 2 bytes to non-volatile memory (\SI{79.0}{\micro\joule})
    \end{itemize}
    Every six hours the device will transmit metered data by doing the following actions:
    \begin{itemize}
    \item Execute \SI{1000000} instructions (\SI{1.14}{\milli\joule})
    \item Transmit 100 bytes of data over ZigBee radio (\SI{12.3}{\milli\joule})
    \item Write 10 bytes to non-volatile memory (\SI{394}{\micro\joule})
    \end{itemize}

@@

    To increase the efficiency of the channel the internal height needs to be reduced by a factor of \SI{52600}, from \SI{71}{\micro\metre} to \SI{1.35}{\nano\metre} (the midpoint between \SI{0.6}{\nano\metre} and \SI{2.1}{\nano\metre}).
    Doing so decreases the normalised output power of the streaming cell to \SI{2.89}{\pico\watt\per\metre}.
    Therefore the width of a channel required to generate the target \SI{5.6}{\milli\watt} is \SI{1.94}{\giga\metre}.

@@

    \section{Conclusion}
    \label{sect:conclusion}
    Approximately one hundred times the required power to operate an automatic meter reader and transmitter is already dissipated inside today's mechanical water meters.
    The streaming cells we constructed produced conversion efficiencies in the order of \SI{1}{\micro\percent}.
    Scaling of these streaming cells to create a viable harvester would result in a \SI{22}{\cubic\metre} volume of glass with a mass of \SI{54}{\mega\gram}.
    This is far from being a practical solution to the battery operated devices currently in use.
